# **IMPORTING DATA**

To begin importing your data onto the cluster, first make sure the files are downloaded on your computer.
Go to Ch-10 (Resources) to download the raw data used in this analysis if you haven't already, or follow along using  your own data! You will also want to download the accompanying metadata files.

## **Getting organized**

Before importing data onto the cluster, it's nice to make sure you have a dedicated directory for all your data files. If you stay organized during your analysis you can pretty much run all your job scripts in the same directory.

Make a directory to work in for this analysis
```{r, eval=FALSE}
mkdir amplicon-analysis

# navigate to the new directory
cd amplicon-analysis

# make 2 more folders called **/rawdata** and **/metadata**
mkdir rawdata
mkdir metadata

#make a file for the bird fecal samples in /rawdata
mkdir rawdata/birds

#make a file for the sediment samples in /rawdata
mkdir rawdata/sediments
```

Check to make sure the new folders are there using the`ls` command

## **Moving data to the cluster** {.tabset}

Transferring data onto the cluster from your local computer desktop can be done using the safe file transfer protocol (SFTP). To do this, first exit the cluster and re-enter the using the following safe file transfer protocol (SFTP) command:

```{r, eval=FALSE}
#first exit the cluster if you are still signed in
exit

#now sign in using the SFTP command
sftp 'username'@graham.computecanada.ca

# Now, move to the amplicon-analysis directory using the 'cd' command
cd projects/'PI-profile'/'username'/amplicon-analysis

```

Next, navigate to the folder on your local computer where the raw data file `raw-data.zip` is stored, usually in a .zip or .gz format. **To run a command in our local computer, always start the command with the letter `l`, like this:**

```{r, eval=FALSE}
#print contents of the rawdata folder on your local computer
lcd <local path to folder containg .FASTQ files>
```

**To move files from your local computer to the cluster:** `put <filename>`
**To move files from the cluster to your local computer:** `get <filename>`

> For example. in order to transfer the files from a local computer into a user directory on the cluster, use the command `put` with a wildcard expression to move all the `.FASTQ.GZ` files in the local directory. To later transfer the QIIME artifacts created in this analysis from the remote computer to our local computer, the `get` command is used.

The `/amplicon-analysis` directory should contain a `/rawdata` folder at this point to move the raw data files into.

```{r, eval=FALSE}
#if you want to transfer the zipped folders into the /rawdata directory on the cluster:
put rawdata-birds.zip rawdata
put rawdata-sediments.zip rawdata

#check to make sure all the files are properly transferred to the cluster
ls rawdata/rawdata-birds
ls rawdata/rawdata-sediments

```

>**NOTE:**
* Use an astericks (*) as a "wildcard" in any regular expression to include multiple files in one command, or avoid writing the full filename over and over again while running commands (see the `put` command below)
* You can check that you're in the right directory using the `pwd` or `lpwd` command

After downloading all of the metadata files from the [Github repo](https://github.com/johannabosch/QIIME2_for_Graham), transfer them over to the cluster like this:

```{r, eval=FALSE}

#transfer the metadata files
lcd <path to metadata files on local computer>
  
cd <path to amplicon-analysis directory>
  
put metadata* metadata/

## OUTPUT:
  
Uploading metadata-birds.csv to /project/6000879/jlb686/amplicon-data/amplicon-analysis/metadata/metadata-birds.csv
metadata-birds.csv                                    100%  585     8.9KB/s   00:00
Uploading metadata-grouped.csv to /project/6000879/jlb686/amplicon-data/amplicon-analysis/metadata/metadata-grouped.csv
metadata-grouped.csv                                  100%  254     5.0KB/s   00:00
Uploading metadata-sediments.csv to /project/6000879/jlb686/amplicon-data/amplicon-analysis/metadata/metadata-sediments.csv
metadata-sediments.csv                                100%  537     4.6KB/s   00:00
Uploading metadata.csv to /project/6000879/jlb686/amplicon-data/amplicon-analysis/metadata/metadata.csv
metadata.csv                                          100%  992    21.2KB/s   00:00

# check to make sure files are there 
ls metadata

## OUTPUT:
metadata-birds.csv      metadata-grouped.csv    metadata-sediments.csv  metadata.csv
```

To proceed to the next steps, exit SFTP, log back into the cluster using `ssh`, and navigate back to your working directory

```{r, eval=FALSE}
exit

ssh jlb686@graham.computecanada.ca

cd projects/'PI-profile'/'username'/amplicon-analysis

```


## **Organizing .FASTQ files**

### Unzip rawdata files

If you need to unzip your compressed rawdata folder on the cluster, use the following command:
```{r, eval=FALSE}
unzip rawdata/<filename>.zip #replace <filename> with your file
```

Unzip the rawdata files for the bird and sediment samples.

If you want to see a portion of what is in one of the `.fastq.gz` rawdata files, you can use the following command:
```{r, eval=FALSE}
zcat rawdata/birds/ATPU01_S125_L001_R1_001.fastq.gz | head
```


Check to make sure the files are in the correct folders using the `ls` command. Here I ran the command so that if you are working with my files you can see what the output should look like:
```{r, eval=FALSE}


ls rawdata/birds

## OUTPUT:

ATPU01_S125_L001_R1_001.fastq.gz  COMU-42_S185_L001_R1_001.fastq.gz
ATPU01_S125_L001_R2_001.fastq.gz  COMU-42_S185_L001_R2_001.fastq.gz
ATPU02_S137_L001_R1_001.fastq.gz  COMU43_S102_L001_R1_001.fastq.gz
ATPU02_S137_L001_R2_001.fastq.gz  COMU43_S102_L001_R2_001.fastq.gz
ATPU03_S149_L001_R1_001.fastq.gz  COMU44_S114_L001_R1_001.fastq.gz
ATPU03_S149_L001_R2_001.fastq.gz  COMU44_S114_L001_R2_001.fastq.gz
ATPU56_S161_L001_R1_001.fastq.gz  COMU46_S126_L001_R1_001.fastq.gz
ATPU56_S161_L001_R2_001.fastq.gz  COMU46_S126_L001_R2_001.fastq.gz
ATPU58_S173_L001_R1_001.fastq.gz  NOGA27_S100_L001_R1_001.fastq.gz
ATPU58_S173_L001_R2_001.fastq.gz  NOGA27_S100_L001_R2_001.fastq.gz
BLKI01_S160_L001_R1_001.fastq.gz  NOGA30_S112_L001_R1_001.fastq.gz
BLKI01_S160_L001_R2_001.fastq.gz  NOGA30_S112_L001_R2_001.fastq.gz
BLKI02_S172_L001_R1_001.fastq.gz  NOGA31_S124_L001_R1_001.fastq.gz
BLKI02_S172_L001_R2_001.fastq.gz  NOGA31_S124_L001_R2_001.fastq.gz
BLKI03_S184_L001_R1_001.fastq.gz  NOGA32_S136_L001_R1_001.fastq.gz
BLKI03_S184_L001_R2_001.fastq.gz  NOGA32_S136_L001_R2_001.fastq.gz
BLKI04_S101_L001_R1_001.fastq.gz  NOGA33_S148_L001_R1_001.fastq.gz
BLKI04_S101_L001_R2_001.fastq.gz  
BLKI05_S113_L001_R1_001.fastq.gz
BLKI05_S113_L001_R2_001.fastq.gz

___

ls rawdata/sediments

## OUTPUT:
FOGO12_S140_L001_R1_001.fastq.gz  REF12_S175_L001_R1_001.fastq.gz
FOGO12_S140_L001_R2_001.fastq.gz  REF12_S175_L001_R2_001.fastq.gz
FOGO1_S187_L001_R1_001.fastq.gz   REF1_S127_L001_R1_001.fastq.gz
FOGO1_S187_L001_R2_001.fastq.gz   REF1_S127_L001_R2_001.fastq.gz
FOGO2_S104_L001_R1_001.fastq.gz   REF2_S139_L001_R1_001.fastq.gz
FOGO2_S104_L001_R2_001.fastq.gz   REF2_S139_L001_R2_001.fastq.gz
FOGO4_S116_L001_R1_001.fastq.gz   REF4_S151_L001_R1_001.fastq.gz
FOGO4_S116_L001_R2_001.fastq.gz   REF4_S151_L001_R2_001.fastq.gz
FOGO6_S128_L001_R1_001.fastq.gz   REF8_S163_L001_R1_001.fastq.gz
FOGO6_S128_L001_R2_001.fastq.gz   REF8_S163_L001_R2_001.fastq.gz
IMP12_S115_L001_R1_001.fastq.gz
IMP12_S115_L001_R2_001.fastq.gz
IMP1_S162_L001_R1_001.fastq.gz
IMP1_S162_L001_R2_001.fastq.gz
IMP2_S174_L001_R1_001.fastq.gz
IMP2_S174_L001_R2_001.fastq.gz
IMP4_S186_L001_R1_001.fastq.gz
IMP4_S186_L001_R2_001.fastq.gz
IMP-8_S103_L001_R1_001.fastq.gz
IMP-8_S103_L001_R2_001.fastq.gz
```
___


&nbsp;
<hr />
<p style="text-align: center;">Github: [johannabosch](https://github.com/johannabosch) </a></p>
<p style="text-align: center;"><span style="color: #808080;"><em>yohannabosch@gmail.com</em></span></p>

<!-- Add icon library -->
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">

<!-- Add font awesome icons -->
<p style="text-align: center;">
    <a href="https://www.instagram.com/yohannabosch/" class="fa fa-instagram"></a>
    <a href="https://www.linkedin.com/in/yan-holtz-2477534a/" class="fa fa-linkedin"></a>
    <a href="https://github.com/johannabosch/" class="fa fa-github"></a>
</p>

&nbsp;